In computing a \emph{task} or sometimes process, is a unit of work that is
executed or will be executed by the processor in a computer or microcontroller.
For many real-time computing systems, the tasks running on them are liable to
meet specific deadlines. I.e.\ the tasks are bound to finish their execution or
respond within a certain time frame. This is especially true for smaller
embedded systems. This might be due to hardware constraints; the system will
not be able to work correctly if the system is over-utilized. Or it could be
critical that certain tasks generate a response in time, e.g.\ a sensor reading
or an actuator. In such scenarios, a task that fails to meet its deadline could
lead to systematic failures.

To make sure tasks adheres to their deadlines, many real-time systems utilizes
a \emph{scheduler} that schedules all tasks. \emph{Scheduling} is the act of
assigning system resources, such as processor time or hardware peripheral
access to the tasks in an orderly fashion. Such that all tasks can finish their
execution in rule with the scheduler's and tasks' constraints.

Many algorithms and policies to schedule tasks currently exists and have been
mathematically proven to be correct.  E.g.\ rate-monotonic
scheduling\cite{ratemonotonic} and earliest deadline first scheduling\cite{edf}
amongst others.  The act of verifying that the system's tasks adhere to the
scheduler policy and are meeting their deadlines is called \emph{schedulability
analysis} or alternatively scheduling analysis.

One way to execute this analysis is to calculate the longest time it takes for
each task to respond, i.e.\ the longest time from when the task is ready to
execute until the end of its execution and compare it to their respective
deadlines.  In order to calculate the worst-case response time (WCRT), the
worst-case execution time (WCET) for the tasks needs to be known. The
difference between WCRT and WCET is that WCET considers the longest execution
time of a task without any interruptions whereas WCRT also considers
interruptions by tasks which have higher priority to
run\cite{hardrealtimecomputingsystems}. If the response time for any task is
calculated to be longer than its deadline, or the system is over-utilized then
the whole system is \emph{unschedulable}.

In this thesis, the tool \emph{RAUK} for automatically verifying an embedded
application's schedulability using a measurement based WCET approach to
caclulate the WCRT of all tasks is presented, inspired by the previous work by
Lindner et al.\ \cite{lindner}. It currently runs on Run-Time Interrupt driven
Concurrency (RTIC) applications, which is a framework for programming real-time
systems for embedded devices in the Rust programming language. It currently
supports all ARM Cortex-M processors.

\section{Background}
Current approaches for automatic WCET calculation on real-time systems usually
utilizes two different methods combined.
\begin{enumerate}
    \item A measurement-based approach, analyzing smaller code sections running
        on target hardware
    \item Static analysis of the program using a model of the system
\end{enumerate}
This is frequently referred as the hybrid approach and is used in safety
critical applications for the aviation industry\cite{rapita} amongst others.
Previous research for the RTFM framework (now called RTIC) showed that it was
possible to utilize the symbolic execution engine KLEE for a measurement-based
WCET analysis for all tasks\cite{lindner}. It showcased the possibilities but
their implementation was no very suitable for practical applications as it
could not analyze or measure WCET on I/O.

When it comes to schedulability analysis tools there are very few active openly
available projects. Most such tools needs the user to manually model the system
in order to verify it, like Cheddar\cite{cheddar} for instance. There have not
been any known attempts for schedulability analysis tools for the RTIC
framework.

\section{Motivation}
For safety-critical systems developed for industries such as automation and
aviation, there is a requirement to show that the systems strictly adheres to
certain timing requirements. In order to carve a path for RTIC and embedded
Rust in general to be used in such systems and industries, the tooling support
for program verification and analysis needs to improve drastically.

Also the lack of easy-to-use tools can be seen as an obstacle for the
open-source community to test their applications. Most tools for schedulability
analysis are framework agnostic but because of that it requires a complete
model of the system which are usually defined in obscure model languages. But a
schedulability tool made specifically for RTIC with both practicality and
easy-of-use in mind, without the need to create a separate model of the system
for testing, could lead to more community-driven projects to be tested more
thoroughly.

\section{Problem definition}
This thesis explores the possibility of developing a tool to automatically
calculate the worst-case response times (WCRT) for tasks in an RTIC application
in order to check if the system is schedulable or not. This by utilizing the
symbolic execution engine KLEE to generate test vectors to target all execution
paths in the application. The assumption is that by targeting all execution
paths in a task, one of the paths will lead to the longest execution. Thus
with the test vectors the WCET of all tasks should be possible to measure.

The thesis also explores what the limitations of such an approach would be, if
any. The goal is to ultimately apply the tool practically on actual working
RTIC applications.

\section{Delimitiations}
The RTIC framework has many features and it would not be possible to test and
analyze all those within the time frame of this thesis. Because of this the
focus have been on verifying tasks which accesses simple resources of primitive
types and offline I/O, i.e.\ I/O where data is written or read at once. Complex
resources such as data structures and accesses of e.g.\ online serial buses
where data is read or written continuously will not be covered.

\section{Thesis structure}
This thesis starts with the introduction to the problem that RAUK solves and
the background for it. Then some brief introduction to related work that has
been made in the field of WCET calculation and schedulability analysis.
Particularly related to embedded Rust. In the theory chapter, the related
theory behind the tools and techniques used will be presented. Then follows
the details of the implementation of RAUK and its main components. After that
the evaluation of the implementation will be presented where the tools is
tested on an RTIC application. Later follows the discussion of the results
in the evaluation and its problems. The thesis ends with the conclusion and
some notes about the future work of this tool.
